---
title: "Assignment-4-EDA"
author: "Johnson_Lee"
date: '2022-07-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(GGally)
library(dplyr)
library(knitr)
```

# About Data
The data used for this analysis has 2479 observation with 8 variables containing data from 2010 -2019 of the NASDAQ returns. The variables are (Year, Lag_1, Lag_2, Lag_3, Lag_4, Lag_5, Volume ,Up_Down). Lag_1 - Lag_5 are the returns from previous day to previous 5 day returns. 

# Importing Data
```{r}
import_data<-read.csv('nasdaq_lagged.csv')
data<-import_data%>%mutate(movement=ifelse(Up_Down=='Up',1,-1))
```

# Checking for Incomplete Data and Duplicates
```{r}
sum(is.na(data))
sum(duplicated(data))
```

# Descriptive Statistics Overall
```{r}
stat <- data %>%
  select(-Year)%>%
  pivot_longer(-Up_Down) %>%
    group_by(Up_Down,name)%>%
  summarise(min = min(value), 
            Q1 = quantile(value, 0.25),
            Median = median(value),
            Average = mean(value),
            Q3 = quantile(value, 0.75),
            Max = max(value),
            SD = sd(value),
            IQR = IQR(value))

stat%>%kable(digits =2)
```

# Descriptive Statistic by Year
```{r}
yearstat <- data %>%
  select(-Up_Down)%>%
  pivot_longer(-Year) %>%
    group_by(Year,name)%>%
  summarise(min = min(value), 
            Q1 = quantile(value, 0.25),
            Median = median(value),
            Average = mean(value),
            Q3 = quantile(value, 0.75),
            Max = max(value),
            SD = sd(value),
            IQR = IQR(value))
```

```{r}
direction<-yearstat%>%filter(name == 'movement')

direction%>%kable(digits =2)
```

```{r}
import_data%>%select(-Year)%>%ggpairs(labs(title= 'Pairs Plot for NASDAQ Return Data'))
```


