---
title: "Final-Project-Baseline"
author: "Johnson_Lee"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(forcats)
library(knitr)
library(yardstick)
library(tidyverse)
library(tidymodels)
library(glmnet)
library(vip)
library(kernlab)
```

# Load Data
```{r}
data<-read.csv('default of credit card clients - adapted.csv')
data<-data%>%mutate_if(is.character, as.factor)
data <- data %>%
    mutate(is_default = fct_relevel(is_default, 'Yes'))
```

# Set Metrics
```{r}
metric <- metric_set(roc_auc, accuracy, precision, recall)
```

# Split Data into Testing / Training
```{r}
set.seed(703412697)
data_split<-initial_split(data, prop = 0.75,strata=is_default)
data_split

data_train<-training(data_split)
data_test<-testing(data_split)
```

# Cross Validation Setup
```{r}
set.seed(64749068)
data_folds <- data_train %>% 
  vfold_cv(v=10, strata=is_default, repeats=3)
```

# Base Recipe (For Logistic Regression)
```{r}
baseline_base_recipe<-
recipe(is_default~., data=data_train)%>%
  step_dummy(gender,educ_level,marital)%>%
  step_zv(all_predictors(),-all_outcomes())%>%
  step_corr(all_predictors(),-all_outcomes())

```

# Enhanced Recipe (Base Recipe + Feature Engineered)
```{r}
baseline_enhanced_recipe<-
recipe(is_default~., data=data_train)%>%
  step_mutate_at(matches('_amount[1-6]'), fn = as.numeric) %>%
    step_mutate(
        bill_amount_credit1 = bill_amount1 < 0,
        bill_amount_credit2 = bill_amount2 < 0,
        bill_amount_credit3 = bill_amount3 < 0,
        bill_amount_credit4 = bill_amount4 < 0,
        bill_amount_credit5 = bill_amount5 < 0,
        bill_amount_credit6 = bill_amount6 < 0
        ) %>%
    step_mutate_at(matches('credit'), fn = as.numeric) %>%
    step_mutate_at(matches('bill_amount[1-6]'), fn = ~ if_else(. < 0, 0, .))%>%
    step_mutate_at(matches('_[0-6]'), fn = as.numeric) %>%
  step_mutate(
    repayment_status0 = pay_0 <0,
    repayment_status2 = pay_2 <0,
    repayment_status3 = pay_3 <0,
    repayment_status4 = pay_4 <0,
    repayment_status5 = pay_5 <0,
    repayment_status6 = pay_6 <0)%>%
  step_mutate_at(matches('status'), fn = as.numeric) %>%
  step_mutate_at(matches('pay[0-6]'), fn = ~ if_else(. < 0, 0, .))%>%
  step_mutate(
    balance1 = bill_amount1-pay_amount1,
    balance2 = bill_amount2-pay_amount2,
    balance3 = bill_amount3-pay_amount3,
    balance4 = bill_amount4-pay_amount4,
    balance5 = bill_amount5-pay_amount5,
  balance6 = bill_amount6-pay_amount6
)%>%
  step_mutate_at(matches('balance'), fn = as.numeric) %>%
  step_mutate_at(matches('bill_amount[1-6]','pay_amount[1-6]'), fn = ~ if_else(. < 0, 0, .))%>%
  step_dummy(gender,educ_level,marital)%>%
  step_zv(all_predictors(),-all_outcomes())%>%
  step_corr(all_predictors(),-all_outcomes())
```

# Setting up Logistic Regression
```{r}
log_reg_model <-
    logistic_reg(penalty=tune(),mixture=1) %>%
    set_engine('glmnet')
```

# Inital Grid
```{r}
set.seed(513116099)
baseline_grid<-grid_latin_hypercube(
  penalty(),size=25
)
```

# Workflows
```{r}
baseline_base_wf<-
  workflow()%>%
  add_recipe(baseline_base_recipe)%>%
  add_model(log_reg_model)
baseline_base_parameters<-extract_parameter_set_dials(baseline_base_wf)
baseline_base_parameters

baseline_enhanced_wf<-
  workflow()%>%
  add_recipe(baseline_enhanced_recipe)%>%
  add_model(log_reg_model)
baseline_enhanced_parameters<-extract_parameter_set_dials(baseline_enhanced_wf)
baseline_enhanced_parameters
```

# First Tune
## Base Recipe
```{r, message=FALSE,warning=FALSE}
baseline_base_tune<-
  baseline_base_wf%>%
  tune_grid(data_folds,
            grid=baseline_grid,
            control = control_resamples(save_pred = TRUE),
            metrics=metric)
```
## Enhanced Recipe
```{r, message=FALSE,warning=FALSE}
baseline_enhanced_tune<-
  baseline_enhanced_wf%>%
  tune_grid(data_folds,
            grid=baseline_grid,
            control = control_resamples(save_pred = TRUE),
            metrics=metric)
```
## Collecting Metrics from Tune Above
```{r}
baseline_base_metric<-
  baseline_base_tune%>%
  collect_metrics()

baseline_enhanced_metric<-
  baseline_enhanced_tune%>%
  collect_metrics()
```

## Visualization of Performance
### Base Recipe Performance
```{r}
baseline_base_metric %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~.metric,scales = 'free_y')
```
```{r}
baseline_base_tune%>%show_best(metric='roc_auc')
```

### Enhanced Recipe Performance
```{r}
baseline_enhanced_metric %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~.metric,scales = 'free_y')
```


```{r}
baseline_enhanced_tune%>%show_best(metric='roc_auc')
```

# Tune Second Model

## Updating Grid Based on Model Outputs Above
```{r}
set.seed(513116099)
baseline_grid_improv<-grid_regular(
  penalty(range= c(-10,0.1)),
  levels=c(25)
  )%>%signif(digits = 10)
```

## Re-Running Tuning With Updated Grid
```{r, message=FALSE,warning=FALSE}
baseline_base_tune_update<-
  baseline_base_wf%>%
  tune_grid(data_folds,
            grid=baseline_grid_improv,
            control = control_resamples(save_pred = TRUE),
            metrics=metric)

baseline_enhanced_tune_update<-
  baseline_enhanced_wf%>%
  tune_grid(data_folds,
            grid=baseline_grid_improv,
            control = control_resamples(save_pred = TRUE),
           metrics=metric)
```

## Collecting Metrics
```{r}
baseline_base_update_metric<-
  baseline_base_tune_update%>%
  collect_metrics()

baseline_enhanced_update_metric<-
  baseline_enhanced_tune_update%>%
  collect_metrics()
```

## Visualization of Performance
### Base Recipe with Updated Grid
```{r}
baseline_base_update_metric %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~.metric,scales = 'free_y')
```

```{r}
 baseline_base_tune_update%>%show_best(metric = 'roc_auc')
```

### Enhanced Recipe with Updated Grid
```{r}
baseline_enhanced_update_metric %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~.metric,scales = 'free_y')
```

```{r}
 baseline_enhanced_tune_update%>%show_best(metric = 'roc_auc')
```

# Best Model's Parameters
```{r}
best_parameters_enhanced_update<-baseline_enhanced_tune_update%>%select_best(metric = 'roc_auc')
```

# Updating Workflow and Fitting Train Data
```{r}
final_workflow_enhanced_update <- finalize_workflow(baseline_enhanced_wf,
                                    parameters=best_parameters_enhanced_update )

final_fit_enhanced_update <- final_workflow_enhanced_update  %>%
    fit(data = data_train)
```

# Creating Predictions on Test Data
```{r}
final_predictions_enhanced_update <- predict(final_fit_enhanced_update, 
                                new_data = data_test)

final_predictions_prob_enhanced_update <- predict(final_fit_enhanced_update, 
                                     new_data = data_test,
                                     type = 'prob')

data_test <- data_test %>%
    bind_cols(final_predictions_enhanced_update, 
              final_predictions_prob_enhanced_update)


```

# Generating Performance Metrics
```{r}

final_metrics_enhanced_update <- data_test %>%
    metrics(truth = is_default, estimate = .pred_class,class1=.pred_Yes)
final_metrics_enhanced_update%>%kable(digits = 4)
```

# Confusion Matrix
```{r}
data_test %>%
    conf_mat(truth = is_default, estimate = .pred_class)
```

# ROC Curve
```{r}
data_test %>%
    roc_curve(truth = is_default,class1=.pred_Yes)%>%
  autoplot
```

# Precision-Recall Curve
```{r}
data_test %>%
    pr_curve(truth = is_default,class1=.pred_Yes)%>%
  autoplot
```

# Saving Important Objects
```{r}
save(data_test,final_workflow_enhanced_update,final_metrics_enhanced_update,
     file = 'baseline_objects.Rda')
```